---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%"
)
```

# irwpkg

`irwpkg` is an R package designed to facilitate easy access to datasets from the Item Response Warehouse ([IRW](https://datapages.github.io/irw/)), offering functions to download, explore, and analyze these datasets. 

## Installation
You can install the development version of `irwpkg` from [GitHub](https://github.com/) with:

```{r}
# install.packages("remotes")
remotes::install_github("hansorlee/irwpkg")
```

## Usage

To use `irwpkg`, load the package using:

```{r}
library(irwpkg)
```

With the package now loaded, we can use the `list_available_datasets()` function to view a list of available datasets. We can then use the `fetch_data(example_data)` function to fetch a dataset named `example_data` from the IRW database into a dataframe.

### View list of datasets
We can also view a list of data sets available for fetching from the IRW using the `list_available_datasets()` function:

```{r}
#| eval: false
# List available datasets
list_available_datasets()
```

### Fetch data
For example, to fetch the `abortion` dataset, we can use:

```{r}
#| eval: false
# Fetch a dataset by name
df <- fetch_data("abortion")
dim(df)
head(df)
```

