---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%"
)
```

# irwpkg

`irwpkg` is an R package designed to facilitate easy access to datasets from the Item Response Warehouse ([IRW](https://datapages.github.io/irw/)), offering functions to download, explore, and analyze these datasets.

## Installation

You can install the development version of `irwpkg` from [GitHub](https://github.com/) with:

```{r}
#| eval: false
# install.packages("remotes")
remotes::install_github("hansorlee/irwpkg")
```


## Usage

To use `irwpkg`, load the package using:

```{r}
library(irwpkg)
```

### Authentication
The IRW datasets are hosted on [Redivis](https://redivis.com), so to access these datasets through `irwpkg`, you’ll need to authenticate with your Redivis account. This connection is managed through the Redivis R client.

**Redivis Authentication Setup**

1. When you first use a function in `irwpkg` that connects to Redivis (e.g. see `list_available_datasets()` below), a browser window will pop up prompting you to sign into Redivis. 

2. After signing in, you’ll be prompted to allow access for the Redivis R Client. Click **Allow**.

3. Once you’ve granted access, close the browser window. You’ll see the message "Authentication was successful" in the R console, and you can then re-run your `irwpkg` function. 

- **Note:** Authentication typically only needs to be done once per session.
- For more details, refer to [Redivis R Client documentation](https://apidocs.redivis.com/client-libraries/redivis-r/getting-started).



### View list of datasets

The list_available_datasets() function provides a list of datasets in the IRW database, showing key metadata (dataset name, row count, and variable count).

```{r}
# List available datasets
list_datasets = list_available_datasets()
dim(list_datasets)
head(list_datasets)
```

### view DB/table metadata

```{r}
db_info = get_database_metadata()

table_info = get_table_metadata("lessR_Mach4")
```



### filter tables by criteria

```{r}
# get tables with a column named "rater"
filter_tables(required_columns="rater")

# get tables with at least 10000 rows and a column named "rater"
filter_tables(n_rows = 10000, required_columns="rater")
```

### Fetch data

Once you have identified a dataset you want, you can use `fetch_data()` to load it as a data frame in R. For example, to fetch the `swmd_mokken` dataset, you can use:

```{r}
# Fetch a dataset by name
swmd_mokken <- fetch_data(name="swmd_mokken")
dim(swmd_mokken)
head(swmd_mokken)
```

### Download data

```{r}
#| eval: false
download_data("swmd_mokken", path = "mydata.csv")
```


### More coming soon :)
