---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%"
)
```

# irwpkg

`irwpkg` is an R package designed to facilitate easy access to datasets from the Item Response Warehouse ([IRW](https://datapages.github.io/irw/)), offering functions to download, explore, and analyze these datasets. 

## Installation
You can install the development version of `irwpkg` from [GitHub](https://github.com/) with:

```{r}
# install.packages("remotes")
remotes::install_github("hansorlee/irwpkg")
```

## Usage

To use `irwpkg`, load the package using:

```{r}
library(irwpkg)
```

With the package now loaded, we can download a dataset using the `fetch_irw()` function or use the `list_available_datasets()` function to view a list of available datasets.

### Download data

For example, to download the `pisa` dataset, we can use:

```{r}
#| eval: false
# Fetch a dataset by name
pisa_df <- fetch_irw(name = "pisa")
names(pisa_df)
```

### View list of datasets

We can also view a list of data sets available for download using the `list_available_datasets()` function:

```{r}
#| eval: false
# List available datasets
list_available_datasets()
```
