---
output: github_document
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  out.width = "100%"
)
```

# irwpkg

`irwpkg` is an R package designed to simplify access to datasets from the Item Response Warehouse ([IRW](https://datapages.github.io/irw/)). It provides tools for exploring, filtering, retrieving, and reformatting datasets, making it easier to analyze item response data.

<!-- ## Installation -->

<!-- To install the development version of `irwpkg` from [GitHub](https://github.com/): -->

<!-- ```{r} -->
<!-- #| eval: false -->
<!-- # Install the remotes package if needed -->
<!-- # install.packages("remotes") -->

<!-- # Install irwpkg from GitHub -->
<!-- remotes::install_github("hansorlee/irwpkg") -->
<!-- ``` -->

## Getting Started

### Loading the Package

After installation, load the `irwpkg` package:

```{r}
#| eval: false
library(irwpkg)
```

### Viewing the Vignette

To help you get started, irwpkg includes a detailed vignette. The vignette demonstrates core functionality, such as:

- Exploring available datasets.

- Visualizing metadata distributions.

- Filtering datasets using custom criteria.

- Fetching datasets into R for analysis.


You can access the vignette using the following commands:

```{r}
#| eval: false
# List all available vignettes for irwpkg
vignette(package = "irwpkg")

# Open the main vignette titled "Report"
vignette("Report", package = "irwpkg")
```

<!-- #### Troubleshooting -->

<!-- If the vignette does not appear, ensure that the package was installed with vignettes enabled. You can do this by reinstalling the package: -->
<!-- ```{r} -->
<!-- #| eval: false -->
<!-- # remotes::install_github("hansorlee/irwpkg", build_vignettes = TRUE) -->
<!-- ``` -->


#### Troubleshooting

##### Missing Vignette

If the vignette does not appear when running `vignette(package = "irwpkg")`, ensure that the package was installed with vignettes enabled. Developers can reinstall the package with the following command:

```{r}
#| eval: false
devtools::install(build_vignettes = TRUE)
```


### Authentication
The IRW datasets are hosted on [Redivis](https://redivis.com), so you’ll need to authenticate with your Redivis account to access these datasets. The authentication process is managed through the Redivis R Client.

**Steps for Authentification**
1. Trigger Authentication

When you first use a function in irwpkg that connects to Redivis (e.g., `list_available_datasets()`), a browser window will open, prompting you to sign in to your Redivis account.

2. Grant Access

After signing in, allow access for the Redivis R Client by clicking **Allow**.

3. Confirmation

Once authentication is successful, close the browser window. You will see the message “Authentication was successful” in the R console.


- **Note:** You only need to authenticate once per session.

- For detailed instructions, refer to the [Redivis R Client documentation](https://apidocs.redivis.com/client-libraries/redivis-r/getting-started).

## Key Features

- Explore Datasets: Quickly browse metadata to understand the structure and range of available datasets.

- Visualize Metadata: Create insightful visualizations to explore key attributes like ID counts, item counts, and sparsity.

- Filter Datasets: Use intuitive criteria to filter datasets based on your research needs.

- Fetch Data: Retrieve individual or multiple datasets for analysis in R.

Please feel free to suggest additional features or report issues on the repository’s GitHub Issues page!
