---
title: "Report"
format: html
vignette: >
  %\VignetteIndexEntry{Report}
  %\VignetteEngine{quarto::html}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(irwpkg)
```

# Introduction

This tutorial demonstrates how to use the irwpkg package to explore, filter, and retrieve data from a structured database of item response datasets. The package includes tools for:

1.  Exploring available datasets and their metadata.

2.  Visualizing distributions of key dataset attributes.

3.  Filtering datasets based on user-defined criteria.

4.  Fetching datasets into R for analysis.

# Explore Available Datasets

## List all available datasets

To view all datasets in the database, use the `list_available_datasets()` function. This provides a summary of dataset names, row counts, and variable counts.

```{r}
#| eval: false
list_datasets <- list_available_datasets()
dim(list_datasets)  # Dimensions of the dataset
head(list_datasets)  # Preview the first few datasets
```

## Get Metadata for the Database or a Specific Table

To get high-level metadata for the entire database:

```{r}
#| eval: false
db_info <- get_database_metadata()
```

To retrieve detailed metadata for a specific table:

```{r}
#| eval: false
table_info <- get_table_metadata("lessR_Mach4")
```

# Understand Metadata Distributions

Before filtering or selecting datasets, it’s helpful to understand the distribution of key dataset attributes.

## View Overall Statistics

The `show_overall_statistics()` function computes key statistics (e.g., range, mean) for attributes such as `id_count`, `item_count`, `resp_count`, and `sparsity`.

NOTE: `sparsity` is calculated using the formula:

$$
\text{sparsity} = \left( \frac{\sqrt{\text{resp_count}}}{\text{id_count}} \right) \times \left( \frac{\sqrt{\text{resp_count}}}{\text{item_count}} \right)
$$

```{r}
#| eval: false
show_overall_statistics()
```

## Visualize Distributions

Use the visualize_metadata_distributions() function to create histograms of dataset attributes. This helps you see the range and frequency of values for attributes like id_count, item_count, resp_count, and sparsity.

```{r}
#| eval: false
visualize_metadata_distributions()
```

You can also focus on specific ranges for individual attributes by specifying ranges:

```{r}
#| eval: false
visualize_metadata_distributions(
  ranges = list(id_count = c(0, 50000), resp_count = c(0, 100000))
)
```

# Filter datasets

After exploring the distributions, you can filter datasets based on specific criteria using the filter_tables() function.

## Filter by Numeric Ranges

Retrieve datasets with attributes that fall within a specified range:

```{r}
#| eval: false
matching_tables <- filter_tables(id_count = c(100, 1000), sparsity = c(0.1, 0.5))
print(matching_tables)
```

## Filter by Presence of Specific Columns

Filter datasets that include specific columns, such as date, rt, or rater:

```{r}
#| eval: false
matching_tables <- filter_tables(has_date = TRUE)

print(matching_tables)
```

## Combine Numeric and Column Presence Criteria

You can combine multiple criteria to refine your search:

```{r}
#| eval: false
matching_tables <- filter_tables(
  id_count = c(100, 10000), 
  has_rater = TRUE
)
print(matching_tables)
```

# Retrieve Datasets

Once you have identified the datasets you need, use the fetch_data() function to load them into R.

## Fetch a Single Dataset

To retrieve a specific dataset:

```{r}
#| eval: false
swmd_mokken <- fetch_data(name = "swmd_mokken")
```

## Fetch Multiple Datasets

You can retrieve multiple datasets at once by providing their names as a character vector:

```{r}
#| eval: false
datasets <- fetch_data(c("fractals_rating", "spelling2pronounce_edwards2023"))
print(names(datasets))
```

Alternatively, you can use the output of filter_tables() directly:

```{r}
#| eval: false
datasets <- fetch_data(matching_tables)
print(names(datasets))
```

# Other utility functions

## download datasets

To save a dataset to a file:

```{r}
#| eval: false
download_data("swmd_mokken", path = "mydata.csv", overwrite=TRUE)
```

# Summary

This tutorial covered the key features of `irwpkg`: 

1. Explore: Use `list_available_datasets()`, `get_database_metadata()`, and `get_table_metadata()` to explore the database.

2.  Visualize: Use `show_overall_statistics()` and `visualize_metadata_distributions()` to understand dataset attributes.

3.  Filter: Use `filter_tables()` to refine your selection based on numeric or column presence criteria.

4.  Retrieve: Use `fetch_data()` to load datasets into R for analysis.
