---
title: "Differential Item Functioning with IRW Data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

A DIF analysis needs two things: a response matrix, and a grouping variable
whose order lines up with the rows of that matrix. The second one is the
awkward part of the IRW workflow, because `irw_long2resp()` returns only `id`,
`item`, and the response -- person covariates are dropped, and some people may
be dropped too. This article uses `irw_covariates()` to keep the two in step,
then runs DIF with `difR` and calibrates items with `irtoys`.

```{r setup, message=FALSE}
library(irw)
library(difR)
library(irtoys)
```

## The data

Many IRW tables carry person covariates in columns prefixed `cov_`, and
`irw_filter()` will find them:

```{r, eval=FALSE}
irw_filter(var = "cov_")
df <- irw_fetch("swmd_mokken")
```

Fetching needs Redivis credentials that this documentation build does not have,
so the article builds a table with a known answer instead. Two groups of 400
people answer the same 10 items; items 3 and 7 are one logit harder for the
focal group and the other eight behave identically. Sixty people, chosen at
random, answered only the first three items -- IRW tables are frequently ragged
like this, and it matters below.

```{r}
a <- rep(1.2, 10)
b_reference <- seq(-1.5, 1.5, length.out = 10)
b_focal <- b_reference
b_focal[c(3, 7)] <- b_focal[c(3, 7)] + 1

reference <- irw_simdata(n_id = 400, n_item = 10, model = "2PL",
                         a = a, b = b_reference, seed = 11)
focal <- irw_simdata(n_id = 400, n_item = 10, model = "2PL",
                     a = a, b = b_focal, seed = 12)
focal$id <- focal$id + 1000

reference$cov_group <- "reference"
focal$cov_group <- "focal"
df <- rbind(reference, focal)

set.seed(9)
quit_early <- sample(unique(df$id), 60)
df <- df[!(df$id %in% quit_early & df$item %in% 4:10), ]
head(df, 3)
```

## Why alignment needs care

Pivot to wide and build the response matrix:

```{r}
wide <- irw_long2resp(df, id_density_threshold = 0.5)
resp <- as.matrix(wide[, setdiff(names(wide), "id"), drop = FALSE])
rownames(resp) <- wide$id
dim(resp)
```

Two things happened. `cov_group` is gone, because `irw_long2resp()` keeps only
the response columns. And the density filter dropped the 60 people who quit
early, so the matrix has fewer rows than the long data has people:

```{r}
length(unique(df$id))
nrow(wide)
```

That is what makes hand-alignment risky. Taking the covariate in its original
order gives a vector of the wrong length:

```{r}
naive <- df$cov_group[!duplicated(df$id)]
length(naive) == nrow(resp)
```

Had the filter dropped nobody, this would have produced a vector of the right
length that was still wrong whenever the wide row order differed from first
appearance -- a silent error rather than a loud one.

`irw_covariates()` reduces the long data to one row per person and, given
`align`, returns those rows in the order of the wide matrix:

```{r}
people <- irw_covariates(df, align = wide)
head(people)
```

```{r}
identical(people$id, wide$id)
table(people$cov_group)
```

```{r}
group <- people$cov_group
```

## Logistic regression DIF

```{r}
dif <- difLogistic(resp, group = group, focal.name = "focal")
colnames(resp)[dif$DIFitems]
```

Statistical significance is easy to reach with hundreds of people per group, so
read the effect sizes rather than the flags:

```{r}
round(dif$deltaR2, 4)
```

Items 3 and 7 -- the two that were built to have DIF -- have much the largest
values. The other flagged items have effect sizes several times smaller, a
reminder that a DIF flag at this sample size is not on its own evidence of a
problem.

## Mantel-Haenszel

A second method on the same inputs, as a cross-check. It flags the same items
here, which is reassuring but not guaranteed in general:

```{r}
mh <- difMH(resp, group = group, focal.name = "focal")
colnames(resp)[mh$DIFitems]
```

## Calibrating items with irtoys

`irtoys` wraps `ltm` and `mirt` behind one interface, which is convenient for
quick estimates and plots:

```{r}
params <- est(resp, model = "2PL", engine = "ltm", nqp = 20)
head(round(params$est, 3))
```

The columns are discrimination, difficulty, and guessing. Calibrating the
pooled sample recovers the generating difficulties reasonably well, except at
the two DIF items, where pooling mixes two different difficulties:

```{r}
data.frame(
  item = colnames(resp),
  b_reference = round(b_reference, 3),
  b_focal = round(b_focal, 3),
  b_pooled_estimate = round(params$est[, 2], 3)
)
```

```{r, fig.alt = "Item characteristic curves for the calibrated two-parameter model."}
plot(irf(params), co = NA, label = TRUE)
```
