---
title: "Fitting IRT Models to IRW Data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

IRW tables arrive in long format with one row per response and the columns
`id`, `item`, and `resp`. Most IRT packages want the opposite shape: a wide
numeric matrix with one row per person and one column per item. This article
covers that handoff, then fits the same data with three packages.

```{r setup, message=FALSE}
library(irw)
library(mirt)
library(ltm)
library(psychotools)
```

## Getting the data

In normal use you fetch a table straight from the warehouse:

```{r, eval=FALSE}
df <- irw_fetch("4thgrade_math_sirt")
```

That call needs Redivis credentials, which this documentation build does not
have. So that these pages show real output, the rest of the article runs on a
simulated table with the same structure, via `irw_simdata()`:

```{r}
df <- irw_simdata(n_id = 500, n_item = 12, model = "2PL", seed = 1)
str(df)
```

## Checking responses before modelling

`irw_check_resp()` flags the two problems that most often break an estimator:
items where everyone gave the same answer, and polytomous categories that are
too thin to estimate.

```{r}
checks <- irw_check_resp(df)
length(checks$single_category_items)
length(checks$sparse_category_items)
```

## Long to wide

`irw_long2resp()` pivots to one row per person. It also applies a default
sparsity filter, dropping people who answered fewer than 10% of items; pass
`id_density_threshold = NULL` to keep everyone.

```{r}
wide <- irw_long2resp(df)
dim(wide)
wide[1:5, 1:5]
```

The result keeps `id` as a column, and its row and column order is not sorted.
IRT packages want just the item columns as a matrix, so drop `id` and move it
to the rownames:

```{r}
resp <- as.matrix(wide[, setdiff(names(wide), "id"), drop = FALSE])
rownames(resp) <- wide$id
dim(resp)
```

That matrix is the input for everything below.

## mirt

```{r, results="hide"}
fit_2pl <- mirt(resp, 1, itemtype = "2PL", verbose = FALSE)
```

```{r}
head(coef(fit_2pl, simplify = TRUE, IRTpars = TRUE)$items)
```

Person ability estimates come back in the row order of `resp`, so they can be
attached to the ids directly:

```{r}
theta <- fscores(fit_2pl)
scores <- data.frame(id = rownames(resp), theta = as.numeric(theta))
head(scores)
```

Comparing nested models is the usual way to justify the extra parameter:

```{r, results="hide"}
fit_1pl <- mirt(resp, 1, itemtype = "Rasch", verbose = FALSE)
```

```{r}
anova(fit_1pl, fit_2pl)
```

## ltm

`ltm` takes a data frame rather than a matrix:

```{r}
fit_rasch <- rasch(as.data.frame(resp))
head(coef(fit_rasch))
```

## psychotools

`psychotools::raschmodel()` is a lean conditional-maximum-likelihood fit, handy
when you want item difficulties without a full IRT apparatus:

```{r}
fit_pc <- raschmodel(resp)
head(itempar(fit_pc))
```

```{r, fig.alt = "Item characteristic curves for the fitted Rasch model."}
plot(fit_pc, type = "curves")
```

## Where to go next

- `vignette("scaling-reliability")` for reliability and Mokken scalability.
- `vignette("dif-linking")` for differential item functioning, which needs
  person covariates carried alongside the response matrix.
