---
title: "Getting Started with irw"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## 📌 Overview
`irw` provides an interface to access datasets from the [Item Response Warehouse (IRW)](https://datapages.github.io/irw/). 

```{r setup, echo=FALSE}
library(irw)
```

---

## Explore Available Datasets
Before fetching data, you can explore the available datasets and their metadata.

List all datasets and their metadata:
```{r, eval=FALSE}
irw_list_tables()
irw_info()
irw_info("4thgrade_math_sirt")
```


---

## Find Datasets by Criteria
Filter datasets based on key attributes.

Example: Filter by respondent count and density range:
```{r, eval=FALSE}
irw_filter(n_participants = c(100, 1000), density = c(0.1, 0.5))
```

Find datasets with rater-based scoring:
```{r, eval=FALSE}
irw_filter(var = "rt")
```

Combine multiple filters:
```{r, eval=FALSE}
irw_filter(n_participants = c(100, 10000), var = "cov_", density = c(0.1, 0.8))
```

---

## Fetch Datasets
Retrieve datasets directly into R for analysis.

Fetch a single dataset:
```{r, eval=FALSE}
swmd_mokken <- irw_fetch("swmd_mokken")
str(swmd_mokken)
```

Fetch multiple datasets:
```{r, eval=FALSE}
datasets <- irw_fetch(c("fractals_rating", "spelling2pronounce_edwards2023"))
print(names(datasets))
```

Fetch datasets that match specific filters:
```{r, eval=FALSE}
matching_tables <- irw_filter(var = "rater", density = NULL)
metadata <- irw_metadata()
metadata[metadata$dataset_name %in% matching_tables, ]

matching_tables <- irw_filter(var = "rater", density = NULL, responses_per_item = c(0, 1000))
datasets <- irw_fetch(matching_tables)
print(names(datasets))
```

---

## Download Datasets
Save datasets locally.

Save as CSV:
```{r, eval=FALSE}
irw_download("gilbert_meta_48")
```

Save dataset metadata as a BibTeX citation:
```{r, eval=FALSE}
irw_save_bibtex("gilbert_meta_48")
```

---


