---
title: "Comparing Models Out of Sample with the IMV"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Fit two IRT models to the same data and the more flexible one always fits
better in sample. The question worth asking is whether it *predicts* better on
responses it has not seen, and by how much. The InterModel Vigorish (IMV)
answers the second half of that: it puts the improvement on the scale of a
weighted coin, so "the 2PL is 3% better than the Rasch model" means something
concrete rather than being a difference of log-likelihoods no one can picture.

Two functions do the work. `irw_predict()` turns a fitted `mirt` model plus a
set of `id`/`item` pairs into predicted probabilities, and `irw_imv()` compares
two such sets of predictions against the observed responses.

```{r setup, message=FALSE}
library(irw)
library(mirt)
```

## What the IMV measures

Each model's predictions are summarized by their geometric mean likelihood,
which is then converted into the weight of a coin with the same entropy. A
model that predicts every response perfectly is a coin that always lands the
same way; a model that predicts nothing is a fair coin. The IMV is the
proportional gain in that weight going from the baseline model to the
comparison model.

Two quick reference points, with no model fitting involved:

```{r}
set.seed(1)
truth <- rbinom(2000, 1, 0.8)

# Same predictions twice: no gain.
irw_imv(truth, p1 = rep(0.8, 2000), p2 = rep(0.8, 2000))

# Knowing the base rate, versus guessing at random.
irw_imv(truth, p1 = rep(0.5, 2000), p2 = rep(0.8, 2000))
```

The statistic is not symmetric — it is the gain from `p1` to `p2`, so swapping
the arguments gives a negative number of a different magnitude, not a sign
flip.

## Getting the data

In normal use you would fetch a table from the warehouse:

```{r, eval=FALSE}
df <- irw_fetch("4thgrade_math_sirt")
```

That needs Redivis credentials, which this documentation build does not have,
so the rest of the article uses `irw_simdata()`. Item discriminations are drawn
with a wide spread, which is exactly the situation where a 2PL should have
something to say that a Rasch model cannot.

```{r}
set.seed(20)
a_true <- rlnorm(12, 0, 0.6)
df <- irw_simdata(n_id = 1000, n_item = 12, model = "2PL", a = a_true, seed = 20)
str(df)
```

## Cross-validating over responses

Hold out a random quarter of the *responses* — not of the people — so that
everyone still appears in the training data and therefore still has an ability
estimate. In each fold, fit both models to the training responses and score the
held-out ones.

```{r}
nfold <- 4
fold <- sample(rep_len(1:nfold, nrow(df)))

omega <- numeric(nfold)

for (k in seq_len(nfold)) {
  train <- df[fold != k, ]
  test  <- df[fold == k, ]

  wide <- irw_long2resp(train, id_density_threshold = NULL)
  items <- wide[, setdiff(names(wide), "id"), drop = FALSE]

  m_rasch <- mirt(items, 1, itemtype = "Rasch", verbose = FALSE)
  m_2pl   <- mirt(items, 1, itemtype = "2PL", verbose = FALSE)

  # Persons need an ability estimate from the training fit to be scored.
  test <- test[test$id %in% wide$id, ]

  z <- irw_predict(m_rasch, wide, newdata = test)
  names(z)[names(z) == "p"] <- "p1"
  z$p2 <- irw_predict(m_2pl, wide, newdata = z[, c("id", "item")])$p

  omega[k] <- irw_imv(z, p1 = "p1", p2 = "p2")
}

round(omega, 4)
mean(omega)
```

The 2PL buys a small but consistent improvement across folds: moving from the
Rasch model to the 2PL is like trading a coin of one weight for a coin roughly
3% more predictable.

Two details in that loop matter more than they look:

- `id_density_threshold = NULL` keeps `irw_long2resp()` from dropping sparse
  respondents. With responses split across folds, everyone looks sparser than
  they are, and a person filtered out of the training frame has no ability
  estimate for the test rows.
- `irw_predict()` is given `wide`, with its `id` column still attached, rather
  than the item-only matrix passed to `mirt()`. That column is how abilities
  from `fscores()` get matched back to people; without it there is nothing to
  stop a silent misalignment between rows and ids.

## Reading the number

The IMV is a *relative* measure and inherits the baseline's units, so it is
only interpretable alongside the comparison being made. A few things to hold on
to:

- **Sign.** Positive means the second model predicts better out of sample.
  Negative is a real and common result: extra parameters that do not
  correspond to anything in the data cost predictive accuracy.
- **Scale.** Values in psychometric model comparisons are typically small.
  A 0.01 gain is not nothing — it is a coin 1% more predictable than the
  baseline's.
- **Out of sample.** Comparing in-sample predictions rewards the more flexible
  model by construction. The cross-validation loop is not optional decoration.

## Reference

Domingue, B. W., Rahal, C., Faul, J., Freese, J., Kanopka, K., Rigos, A.,
Stenhaug, B., & Tripathi, A. (2021). InterModel Vigorish (IMV): A novel
approach for quantifying predictive accuracy with binary outcomes.
<https://doi.org/10.31235/osf.io/8sgz5>
