---
title: "Reliability and Scalability with IRW Data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Before fitting a measurement model it is worth asking whether the items hang
together at all. This article covers internal consistency with `psych` and
nonparametric scalability with `mokken`, using a real IRW table that ships with
the package.

```{r setup, message=FALSE}
library(irw)
library(psych)
library(mokken)
```

## The data

`swmd_mokken` is bundled with `irw`, so this article runs on genuine warehouse
data with no network access. Fetching it yourself would look like this:

```{r, eval=FALSE}
df <- irw_fetch("swmd_mokken")
```

```{r}
df <- swmd_mokken
str(df)
```

Note the `rater` column: each person-item pair was scored by several raters, so
there are many more rows than the 30 x 7 = 210 person-item combinations.

```{r}
nrow(df)
length(unique(df$id)) * length(unique(df$item))
```

## Collapsing raters

`irw_long2resp()` aggregates duplicate `id`-`item` pairs, and how it does so
matters here. The default for a numeric response is `"mean"`, which produces
fractional scores that neither package below will accept as ordinal
categories. `"mode"` keeps the responses on their original 1-5 scale by taking
each rater panel's most common score:

```{r}
wide <- irw_long2resp(df, agg_method = "mode", id_density_threshold = NULL)
dim(wide)
wide[1:3, ]
```

```{r}
resp <- as.matrix(wide[, setdiff(names(wide), "id"), drop = FALSE])
rownames(resp) <- wide$id
range(resp)
```

## Internal consistency

```{r}
alpha(resp)$total
```

```{r}
omega(resp, nfactors = 1, plot = FALSE)$omega.tot
```

## Mokken scalability

Scalability coefficients ask whether the items form a cumulative scale.
Guttman's rule of thumb reads H below 0.3 as unscalable, 0.3-0.4 as weak,
0.4-0.5 as moderate, and above 0.5 as strong.

```{r}
H <- coefH(as.data.frame(resp))
H$Hi
H$H
```

Every item clears 0.5 and the scale coefficient is comfortably strong.

Automated item selection confirms that the items belong to a single scale --
every item is assigned to scale 1 at the default lower bound of 0.3:

```{r}
aisp(as.data.frame(resp))
```

`aisp()` warns that the items show varying numbers of observed response
categories. With only 30 respondents some of the five categories simply never
came up for some items. That is a property of this small table rather than a
mistake, but on a larger table it is worth checking with `irw_check_resp()`,
which reports sparse categories item by item:

```{r}
checks <- irw_check_resp(df)
names(checks$sparse_category_items)
```

## Monotone homogeneity

The monotonicity check needs enough people per rest-score group to estimate
item-step response functions. With only 30 respondents, `check.monotonicity()`
stops at its default group size, so the group size has to be set by hand:

```{r}
mono <- check.monotonicity(as.data.frame(resp), minsize = 10)
summary(mono)
```

No violations, but treat that as illustrative rather than conclusive. The
`#ac` column counts the comparisons the check was actually able to make, and at
this sample size there are very few of them: `minsize = 5` yields 18 active
comparisons and 3 violations, `minsize = 10` yields 3 and 0, and `minsize = 15`
yields 1 and 0. A clean result from three comparisons is weak evidence. A
serious monotonicity assessment wants a larger table; `irw_filter()` will find
one:

```{r, eval=FALSE}
irw_filter(n_participants = c(500, 5000), density = c(0.8, 1))
```
