---
title: "SEM and Multilevel Models with IRW Data"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Not every measurement model wants a response matrix. This article fits a
categorical confirmatory factor analysis with `lavaan`, which needs the wide
form, and then a Rasch model as a generalized linear mixed model with `lme4`,
which consumes IRW long format directly with no reshape at all.

```{r setup, message=FALSE}
library(irw)
library(lavaan)
library(lme4)
```

## The data

```{r, eval=FALSE}
df <- irw_fetch("4thgrade_math_sirt")
```

That call needs Redivis credentials, which this documentation build does not
have, so the article runs on a simulated table of the same shape:

```{r}
df <- irw_simdata(n_id = 500, n_item = 8, model = "2PL", seed = 3)
head(df, 3)
```

## Categorical CFA with lavaan

`lavaan` needs one column per item, so pivot first:

```{r}
wide <- irw_long2resp(df)
resp <- as.data.frame(wide[, setdiff(names(wide), "id"), drop = FALSE])
```

Because the item columns are already named `item_1` ... `item_8`, the model
syntax can be built from `names(resp)` rather than typed out:

```{r}
items <- names(resp)
model <- paste0("F =~ ", paste(items, collapse = " + "))
model
```

Binary items must be declared `ordered`, which switches `lavaan` to a
diagonally weighted least squares estimator with a probit link. A CFA fitted
this way is equivalent to a two-parameter normal-ogive IRT model:

```{r}
fit <- cfa(model, data = resp, ordered = items)
```

Declaring the items ordered switches the estimator to DWLS with a scaled and
shifted test statistic, so the `.scaled` fit measures are the ones to read;
the unscaled `pvalue` is reported as `NA` under this test:

```{r}
fitMeasures(fit, c("chisq.scaled", "df.scaled", "pvalue.scaled",
                   "cfi.scaled", "tli.scaled", "rmsea.scaled"))
```

The fit is close to perfect, which is what simulated unidimensional data
should produce. Real IRW tables rarely look this tidy.

```{r}
head(parameterEstimates(fit, standardized = TRUE)[, c("lhs", "op", "rhs", "est", "std.all")], 8)
```

The standardized loadings are a monotone transformation of IRT
discriminations, so items that load highly here are the same items `mirt`
would give a high `a` parameter.

## Rasch as a mixed model with lme4

This is the one workflow where IRW's native shape is exactly what the package
wants. A Rasch model is a logistic mixed model with a random person intercept
and fixed item effects, so the long data frame goes straight in:

```{r}
df$id <- factor(df$id)
df$item <- factor(df$item)
```

```{r}
fit_glmm <- glmer(resp ~ 0 + item + (1 | id), data = df, family = binomial)
```

The fixed effects are item easiness, so negate them to read them as the item
difficulties an IRT package would report:

```{r}
difficulty <- -fixef(fit_glmm)
head(round(difficulty, 3))
```

The person variance is the squared latent trait standard deviation:

```{r}
VarCorr(fit_glmm)
```

### Why bother

Writing the Rasch model this way buys you everything `lme4` can do. Person
covariates enter as fixed effects, giving an explanatory IRT model in one line,
and crossed random effects handle raters without a separate package. IRW tables
that carry a `rater` column suit the latter directly:

```{r, eval=FALSE}
# for a table with a rater column, e.g. irw_fetch("swmd_mokken")
glmer(resp ~ 0 + item + (1 | id) + (1 | rater), data = df, family = binomial)
```

See `vignette("dif-linking")` for adding a person covariate, which is where
`irw_covariates()` comes in.
